Social Attention for Autonomous Decision-Making in Dense Traffic

### 2019-09-04 main.tex
Update content

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
 \usepackage[preprint]{neurips_2019_ml4ad}

@@

% Custom packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{float}
\usetikzlibrary{arrows,automata}
\usepackage{pgfplots}
\usetikzlibrary{intersections}

@@

\author{%
  Edouard Leurent\thanks{Equal contribution.} \\
  Inria, SequeL team\\
  Renault Group\\
  Paris, France\\
  \texttt{edouard.leurent@renault.com} \\
  % examples of more authors
   \And
  Jean Mercat$^*$ \\
  Laboratoire des signaux et des syst\`emes,
  Centrale-Sup\'elec\\
  Renault Group\\
  Paris, France\\
  \texttt{jean.mercat@renault.com} \\
}

@@

@@

\subsection{Model-Free Deep Reinforcement Learning}

- MDP
- Value function
- DQN

\subsection{State Representation for Social Interactions}
In the state:
Kinematics
Not fixed size
Occupancy grid

@@

Weight sharing ?

@@

	\draw (DEC1.east) -- (Y1.west);

	%        \draw[decorate,decoration={brace,mirror}] (X1.north west) -- node[left=10pt] {$X$} (X3.south west);
	\end{tikzpicture}
	\caption{Block diagram of our model architecture.}
	\label{sch_whole_model}

@@

	\coordinate[right of=LQ1, node distance=2.6cm](Q1){};
	\node[below of=Q1, node distance=1.5cm](Q2){$Q = \mathbf{q}_0$};
	\node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_0$};
	\node[right of=LV1, node distance=1cm](V1){$\mathbf{v}_0$};

	\draw (LQ1) -- (Q1);
	\draw (Q1) -- (Q2);
	\draw (LK1) -- (K1);
	\draw (LV1) -- (V1);

@@

	\coordinate[right of=K1, node distance=0.3cm](TOP){};
	\coordinate[right of=V3, node distance=0.3cm](BOT){};
	\draw[decorate,decoration={brace}] (TOP) -- node[left=5pt]{} (BOT);

	\node[right of=X2, text width=3cm, node distance=5.5cm, below=-1.0cm](EQ){\footnotesize \[K = \left( \begin{matrix}
		\mathbf{k}_0 \\
		\vdots \\
		\mathbf{k}_{n}
		\end{matrix} \right)\]
		\\
		\footnotesize \[ V = \left( \begin{matrix}
		\mathbf{v}_0 \\
		\vdots \\
		\mathbf{v}_{n}
		\end{matrix}\right) \]
	};

@@

\section{Experiments}
\subsection{The intersection task}

We consider a situation where interactions between vehicles is important: crossing an unsignalized intersection.

To that end, we use the highway-env environment and propose a novel intersection task.

- The intersection: cross (North/South/East/West), single lane or 2 lane ?
- The other vehicles: are spawned randomly, and plan a route to a random direction.
Follow a Kinematic Bicycle Model, lane keeping with a low-level steering controller, longitudinal control with IDM

\subsection{The baselines}
\subsection{Performances}
\subsection{Attention interpretation}

@@

>>>>

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
 \usepackage[preprint, nonatbib]{neurips_2019_ml4ad}

@@

% Custom packages
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{float}
\usetikzlibrary{arrows,automata}
\usepackage{pgfplots}
\usetikzlibrary{intersections}
\usepackage{subcaption}

\input{mathdef}

@@

\author{%
  Edouard Leurent\thanks{Equal contribution.} \\
  SequeL team, INRIA Lille -- Nord Europe\\
  Renault Group, France\\
  \texttt{edouard.leurent@inria.fr} \\
  % examples of more authors
   \And
  Jean Mercat$^*$ \\
  Laboratoire des signaux et des syst\`emes, Centrale-Sup\'elec\\
  Renault Group, France\\
  \texttt{jean.mercat@renault.com} \\
}

@@

Contributions:
- We propose an attention-based architecture for decision-making involving social interactions, that accounts for two-way interactions between the ego-vehicle and every other traffic participant
- We evaluate our model on a challenging intersection-crossing task involving up to 20 vehicles perceived simultaneously.

@@

\paragraph{Model-free deep reinforcement learning} Reinforcement Learning (RL) is a general framework for decision-making under uncertainty. It frames the learning objective as the optimal control of a Markov Decision Process  $(S, A, P, R, \gamma)$ with measurable state space $S$, action space $A$, unknown reward function $R\in\Real^{S \times A}$, and unknown dynamics $P\in \cM(S)^{S \times A}$, where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$. We seek a policy $\pi\in\cM(A)^S$ with maximal expected $\gamma$-discounted cumulative reward, called the value function $V^\pi$. Formally,

\begin{align*}
V^\pi(s) &\eqdef \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\condbar s_0=s, a_t\sim \pi(a_t|s_t), s_{t+1}\sim P(s_{t+1}|s_t, a_t)\right]\\
Q^\pi(s, a) &\eqdef R(s, a) + \gamma \expectedvalueover{s'\sim P(s'|s, a)} V^\pi(s')
\end{align*}

The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation:
\begin{equation*}
Q^*(s, a) = (\cT Q^*) (s, a) \eqdef \expectedvalueover{s'\sim P(s'|s, a)} \max_{a'\in A} \left[R(s, a) + \gamma Q^*(s', a')\right]
\end{equation*}

As $Q^*$ is a fixed-point of the Bellman Operator $\cT$ -- which is a contraction --, it can be computed by applying $\cT$ iteratively in a fixed-point iteration fashion. The \emph{Q-learning} algorithm \citep{Watkins} follows this procedure by applying a sampling version $\cT$ to a batch of collected experience. When dealing with a continuous state space $S$, we need to employ function approximation in order to generalise to nearby states. The \emph{Deep Q-Network} (DQN) algorithm \citep{Mnih} implements this idea by using a neural network model to represent the action-value function $Q$.

\paragraph{State-representation for social interactions}{

In the state:
Kinematics
Not fixed size
Occupancy grid

@@

Weight sharing ?
Socially aware}

\paragraph{Attention mechanisms} {in neural networks have been able to introduce inter-dependencies within a variable number of inputs.
It has been used for pedestrian trajectory forecasting in~\cite{Vemula2018} with spatiotemporal graphs and
in~\cite{Sadeghian2019CVPR} with spatial and social attention using a generative neural network.
In~\cite{Sadeghian2018ECCV}, attention over top-view road scene images for car trajectory forecasting is used.
Multi-head attention mechanism has been developed in~\cite{Vaswani2017} for sentence translation.
In~\cite{Messaoud2019} a mechanism called non-local multi-head attention is developed.
However, this is a spatial attention that does not allow vehicle-to-vehicle attention.

In the present work, we use a multi-head social attention mechanism working together with
recurrent neural networks to forecast in the form of sequences of multi-modal position probability density
functions.}

@@

	\draw (DEC1.east) -- (Y1.west);
	\end{tikzpicture}
	\caption{Block diagram of our model architecture.}
	\label{sch_whole_model}

@@

	\node[right of=LQ1, node distance=1cm](Q1){$\mathbf{q}_0$};
	\node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_0$};
	\node[right of=LV1, node distance=1cm](V1){$\mathbf{v}_0$};

	\draw (LQ1) -- (Q1);
	\draw (LK1) -- (K1);
	\draw (LV1) -- (V1);

@@

	\coordinate[right of=Q1, node distance=0.3cm](TOP){};
	\coordinate[right of=V3, node distance=0.3cm](BOT){};
	\draw[decorate,decoration={brace}] (TOP) -- node[left=5pt]{} (BOT);

	\node[right of=X2, text width=3cm, node distance=5.5cm](EQ){
		\footnotesize \[Q = \left( \begin{matrix}
		\mathbf{q}_0
		\end{matrix} \right)\]
		\\
		\footnotesize \[K = \left( \begin{matrix}
		\mathbf{k}_0 \\
		\vdots \\
		\mathbf{k}_{n}
		\end{matrix} \right)\]
		\\
		\footnotesize \[ V = \left( \begin{matrix}
		\mathbf{v}_0 \\
		\vdots \\
		\mathbf{v}_{n}
		\end{matrix}\right) \]
	};

@@

\section{Experiments}
\subsection{Description of the environment}

In this application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{highway-env} for simulated highway driving and behavioural decision-making. We propose a new task where vehicles must interact significantly with each others: a four-way intersection. The scene, composed of two roads crossing perpendicularly, is populated with several traffic participants initialized with random positions, velocities, and destinations. As in \citep{highway-env}, these vehicle are simulated with the Kinematic Bicycle Model, their lateral control is achieved by a low-level steering controller and their longitudinal behaviour follows the Intelligent Driver Model. However, this model only considers same-lane interactions and special care was required to prevent the lateral collisions at the intersection. To that end, we implemented the following simplistic behaviour: each vehicle predicts the future positions of its neighbours over a three-second horizon by using a constant velocity model. In case of predicted collision with a neighbour, the yielding vehicle is determined randomly and must brake until the collision prediction ceases.

In this context, the agent must drive a vehicle by controlling its acceleration chosen from a finite set of actions $A = \{\texttt{SLOWER}, \texttt{NO-OP}, \texttt{FASTER}\}$. The lateral control is performed automatically by a low-level controller, such that the problem complexity is focused on the high-level interactions with other vehicles, namely the decision to either yield or take way. The agent is rewarded by $1$ when it drives at maximum velocity, $0$ otherwise, and by $-5$ when a collision occurs.

\subsection{Agents}

\subsection{Results}

\begin{table}
	\centering
	\begin{tabular}{lcc}
		\toprule
		Architecture & MLP & Ego-Attention \\
		\midrule 
		Layers sizes & [128, 128] & [64, 64], 64, [64, 64] \\
		Number of parameters & \textbf{3.0e4} & 3.4e4 \\
		Training Time (w.r.t. MLP) & \textbf{1} & 1.25 \\
		Mean total reward & 3.6 & \textbf{6.4} \\ 
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}
	\centering
	\begin{subfigure}[t]{.65\linewidth}
		\centering\includegraphics[width=\linewidth]{img/total_rewards}
		\caption{Cumulative reward}
	\end{subfigure}
\\
	\begin{subfigure}[t]{.65\linewidth}
		\centering\includegraphics[width=\linewidth]{img/total_length}
		\caption{Episode length}
	\end{subfigure}
\\
	\begin{subfigure}[t]{.65\linewidth}
		\centering\includegraphics[width=\linewidth]{img/collision}
		\caption{Probability of collision}
	\end{subfigure}
\end{figure}

\subsection{Attention interpretation}

@@

\bibliographystyle{named}
\bibliography{references}
